---
title: "Exploration into Gapminder"
author: "Nicole Hawe"
date: "September 25, 2017"
output: github_document
---

Introduction
------------

Within this second homework assignment, we explore the Gapminder dataset in more detail. 


This page will cover the work flow for: 

- Overall Properties of the Gapminder Dataset

- Variable Property Exploration

- Data Visualization using ggplot2

- Data Filtering/ Piping

- Reflections


Overall Properties of the Dataset
---------------------------------

The first step is to load the Gapminder dataset and packages we need!

```{r}
library(gapminder)
library(tidyverse)
library(dplyr)

```

So the assignment begins with the question: what is Gapminder? Is it a data.frame, a matrix, a vector, a list? Well, I decided to check this by using typeof() function.

```{r}
typeof(gapminder)
```

However I would like to know the class of gapminder

```{r}
class(gapminder)
```

It appears to be a data frame, or more specifically a tibble!

Next, information regarding the struture of the dataset was asked such as:

- How many variables/columns do we have? We can use ncol in R

- How many rows/observations do we have? We can use nrow in R


```{r}
ncol(gapminder)
```

```{r}
nrow(gapminder)
```
So it looks like this dataset has 6 columns with 1704 rows!

The it was asked whether we can get these facts about “extent” or “size” in more than one way? Can you imagine different functions being useful in different contexts?

we can find the length by

```{r}
length(gapminder)
```

Or the dimensions by

```{r}
dim(gapminder)
```

We can also find the names of all th columns by

```{r}
names(gapminder)
```

Or get an overall sense of the structure using
```{r}
str(gapminder)
```

The str function is a way of viewing the number of columns and rows along with the class and name of each variable! This is a great way to see many overall facts about the dataset!

```{r}
summary(gapminder)
```


The summary function shows the dataset in a different way and for variables that are numbers or integers it shows importants statistics on them!

As you can see this information can be found in several ways!

Next we needed to find the type of each variable in this dataset using the typeof function

```{r}
typeof(gapminder$country)
typeof(gapminder$continent)
typeof(gapminder$year)
typeof(gapminder$lifeExp)
typeof(gapminder$pop)
typeof(gapminder$gdpPercap)
```

Exploration of Individual Variables
-----------------------------------

For this part of the homework assignment we were to pick at least one categorical variable and at least one quantitative variable to explore.

For my categorical variable I chose **Year**

For my quantitative variable I chose **LifeExp**

For these we needed to find the possible values (or range, whichever is appropriate) of each variable?

What values are typical? What’s the spread? What’s the distribution? 

So beginning with Year:

I wanted a nice summary to start with 
```{r}
summary(gapminder$year)
```

```{r}
table(gapminder$year)
```

From this quick summary and table I see that the minimum year is 1952 and it ranges to the maximum being 2007. Both the median and mean are 1980. I also appears that samples were taken every 5 years starting in 1952.

Moving onto the quantitative variable LifeExp:

```{r}
summary(gapminder$lifeExp)
```
 
From this summary we can see the minimum value is 23.60 whereas the max is 82.60. This is quite a large spread. The median lies at 60.71 and the mean at 59.47.

For a value such as LifeExp a table may be too large and unuseful.

A quick plot can sometimes help though!

```{r}
hist(gapminder$lifeExp)
```


Gapminder Dataset Visualization Using ggplot2
---------------------------------------------

In this part of the assignment we were to make a few plots, using the same variable we chose to characterize numerically. The goal was to try to explore more than one plot type. For example:

- A scatterplot of two quantitative variables.

- A plot of one quantitative variable. Maybe a histogram or densityplot or frequency polygon.

- A plot of one quantitative variable and one categorical. Maybe boxplots for several continents or countries.

For a scatterplot I chose to see the relationship between LifeExp and gdpPerCap

```{r}
ggplot(gapminder, aes(x = gdpPercap, y = lifeExp)) + geom_point()
```

Next I wanted to check out a histogram for gdpPercap

```{r}
hist(gapminder$gdpPercap)
```

ALso just to refresh your memory a histogram of LifeExp

```{r}
hist(gapminder$lifeExp)
```

Now I wanted to test out a plot while changing the colour:

```{r}
ggplot(gapminder, aes(x=year, y=lifeExp)) +
  geom_point(size=2, aes(colour = continent))
```


Next I wanted to try something a little more interesting. I wanted to look at LifeExp versus Year, but I wanted to divide this into countries. I arbitrarily picked Canda, Afghanistan, Sweden, Rwanda, and Spain. Let's see if we can pull this off.

Please note I used [This helpful link](https://github.com/jennybc/ggplot2-tutorial)


```{r}
jCountries <- c("Canada", "Afghanistan", "Sweden", "Rwanda", "Spain")
ggplot(subset(gapminder, country %in% jCountries), aes(x = year, y = lifeExp, color = country)) + geom_line() + geom_point()
```

You can really see the large separation between more developed countries (Sweden, Canada, and Spain) and lesser like Rwanada and afghanistan. Also you see a severe dip in LifeExp in Rwanada around 1992, I'm guessing due to the war happening around that time.


Piping Practice with Dyplyr Function
------------------------------------

Next we went on to practice the filter function as well as piping the dyplyr filter and select functions together! These functions help narrow down and focus in on the data we care about. 

To start I wanted to find the mean lifeExp of each continent:

```{r}
gapminder %>%
  group_by(continent) %>%
  summarize(meanLife = round(mean(lifeExp), 1)) %>%
  arrange(desc(meanLife))
```

Here I get a table that has filtered the gapminder dataset to show the mean lifeExp (meanLife) of each continent in descending order based on the meanLife value!

Next maybe I want to look at the variables but only for specific countries in a specific year:

```{r}
gapminder %>%
  filter(country %in% c("Sweden", "Norway", "Switzerland"), year == 2002)
```

Now What if we tried incorporating these dyplyr function with ggplot!


Say I only wanted the LifeExp versus year for Canada I could filter this:

```{r}
ggplot(gapminder %>% 
         filter(country == "Canada"), aes(x = year, y = lifeExp)) + geom_line() + geom_point()
```


Next I wanted to attempt to use filter and select together to plot LifeExp versus Year for Cambodia:

```{r}
filter(gapminder, country == c("Cambodia")) %>%
  select(year, lifeExp) %>%
  ggplot(aes(year, lifeExp)) + geom_line() + geom_point(colour="magenta")
```

Again you see a drastic drop in LifeExp around 1975, again around the time this country was experiencing a civil war.

Piping in a very powerful way to string together many functions.

#### We Survived!

Reflections on This Assignment
------------------------------

Due to be a serious newcomer to this world of Stats I decided to stop this assignment here and not dive into the "I want to do more" section.

This assignment challenged me a lot and I spent a long time sifting through all of the tutorials to wrap my head around the proper use of different functions.

I feel as though it was a big step up from Assignment 1, but this may just be for me!

It took me a long time to figure out the order of things and to be honest I would love to hear your feedback on my ordering.

I enjoyed working with ggplot2 a lot because it is a great visual tool an could see how the different functions altered the graph. This helped me understand a lot.
