---
title: "But I Want To Do More"
author: "Nicole Hawe"
date: "October 3, 2017"
output: github_document
---

```{r}
library(gapminder)
library(tidyverse)
library(knitr)
library(kableExtra)
library(gridExtra)
# loaded grid extra to help put table and figure side by side
```

#### Layout Stretch Goal

Okay so I had trouble using the code found [here](http://gist.github.com/jennybc/e9e9aba6ba18c72cec26)
So I found another way using the package [gridExtra](https://cran.r-project.org/web/packages/gridExtra/gridExtra.pdf)



So remember from Task 3 we had:
```{r}
n5 <- gapminder %>% 
  group_by(year) %>% 
  summarise(trim_mean_lifeExp = mean(lifeExp, trim=0.25)) %>% 
  arrange(year) %>% 
  select(year, trim_mean_lifeExp)

t1 <- kable(n5)
t1

```
Also the accompanying plot:
```{r}
p2 <- ggplot(n5, aes(x=year, y=trim_mean_lifeExp)) +
  geom_point(colour="purple", size=3) +
  labs(x="Year", y="Trimmed Mean LifeExp")+
  ggtitle("Trimmed Mean Life Expectancy") +
  theme(plot.title = element_text(hjust=0.5))
p2
```


Now in order to get them side by side we need to define a tableGrob for gridExtra:

```{r}
t1a <- tableGrob(t1)
grid.arrange(p2, t1a, nrow=1)
```

This does not look fantastic but is better than nothing. I'm looking forward to hearing other suggestions!

#### Table Stretch Goals

I decided to venture into pander to make a different style of table:

```{r}

library(pander)
```

```{r}
pander(n5)
```
There this gives another table making option, however I didn't go too crazy with customization, I'll leave that for the future.

I chose pander with the help found [here](https://twitter.com/polesasunder/status/464132152347475968)

Back to My [HW3 Folder](https://github.com/nicolehawe/STAT545-HW-Hawe-Nicole/tree/master/HW03)

