---
title: "Task 2"
author: "Nicole Hawe"
date: "October 2, 2017"
output: github_document
---

```{r}
library(gapminder)
library(tidyverse)
library(knitr)
library(kableExtra)
```


#### Look at the Spread of GDP per Capita Within the Continents

I was a little bit unclear on what was wanted with this question so to begin I used all of the data to plot the complete spread of GdpPerCap for all the continents: 

```{r}
gapminder %>%
  group_by(continent) %>%
  ggplot(aes(gdpPercap, fill=continent)) + 
  geom_histogram(color="black") +
  facet_wrap(~continent, scales="free")
```

Then I created a table however due to the size of a table for this sort of data I decided to only show part of it using the "head" function:

```{r}
t3 <- gapminder %>%
  group_by(continent) %>%
  select(continent,gdpPercap)

t4 <- head(t3)
kable(t4)
  

```


I could also interpret the spread to mean the mean GdpPerCap with the standard deviation for each continent so I did that as well:

```{r}
n4 <- gapminder %>% 
  group_by(continent) %>% 
  summarize(mean_gdp=mean(gdpPercap),sd_gdp=sd(gdpPercap))
kable(n4,format='markdown')
```

This data can then be visulaized with a box plot:

```{r}
gapminder %>%
  ggplot(aes(x=continent, y=gdpPercap)) + 
  geom_boxplot(aes(group=continent, colour=continent, fill=continent),alpha=0.5)+
  ggtitle("GDP Per Capita Per Continent")+
  theme(plot.title = element_text(hjust = 0.5))
  
```

